---
title: "Founders Circadian Rhythm Fibroblast Data: Heritability and Plotting"
author: "Michael C. Saul (michael.saul [at] jax.org)"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: false
      smooth_scroll: false
    number_sections: true
    fig_retina: true
    self_contained: true
    code_folding: show
    theme: flatly
    highlight: default
    df_print: paged
---

# Background

## Motivation

### Purpose of analysis

This analysis is performed to process the circadian rhythm amplitude data and calculate:

1) Heritability
2) Strain-by-sex means and standard errors for plotting

**These data were collected by Sam Moon-Kim.** The preprocessing script `circadian_fibroblastmport_preprocess.Rmd` was used to filter the data to only those data requiring reporting.

# Analysis

## Setup

### Working directory

All chunks will run in the working folder for this experiment.

```{r setup}
script_root_dir = "~/Box/manuscripts/founders_heritability/"
knitr::opts_knit$set(root.dir = script_root_dir)
```

Getting current date

```{r}
current_date = format(Sys.time(),"%Y-%m-%d")
```

### R libraries

Calling R libraries necessary for this analysis.

```{r}
library("tidyverse")
library("lubridate")
library("cowplot")
library("MASS")
library("ggbeeswarm")
```

## Importing data

## Importing and displaying raw fibroblast amplitude data

Getting data from the filtering script.

```{r}
fibroblast_amplitude = readRDS(paste(script_root_dir, "data/circadian_amplitude_data_preprocessed_2019-09-04.RDS", sep = ""))
```

Displaying raw data.

```{r}
fibroblast_amplitude
```

Getting the amplitude means and using these for heritability calculations.

```{r}
fibroblast_amplitude_mean = fibroblast_amplitude %>%
  group_by(Subject) %>%
  summarize(Strain = unique(Strain),
            Sex = unique(Sex),
            Cohort = unique(Cohort),
            Amplitude_Mean = mean(Amplitude))
fibroblast_amplitude_mean = as.data.frame(fibroblast_amplitude_mean)
```

Getting founders strain key

```{r}
founders_key = readRDS("~/Box/CC_DO_founders/cc_do_founders_key.RDS")
row.names(founders_key) = founders_key$strain
```

### Fitting ANOVA model for fibroblast data

Looking at the residuals of the simplest model first.

```{r}
fibroblast_amplitude_resid = fibroblast_amplitude_mean
fibroblast_amplitude_resid$Amplitude_Resid = resid(lm(Amplitude_Mean ~ Strain, data = fibroblast_amplitude_resid))
fibroblast_amplitude_resid$Strain = factor(fibroblast_amplitude_resid$Strain,
                                           levels = founders_key$strain,
                                           ordered = TRUE)

ggplot(data = fibroblast_amplitude_resid, aes(x = Strain, y = Amplitude_Resid, color = Strain, shape = Sex)) +
  geom_hline(yintercept = 0, col = "#999999") +
  geom_beeswarm(size = 2) + 
  scale_color_manual(values = founders_key$collaborative_cross_color_broman) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_shape_manual(values = c(16,23))
```

The residuals indicate heteroscedasticity -- especially comparing 129 to NOD. Doing Box-Cox to see if a transform will help.

```{r}
boxcox(lm(Amplitude_Mean ~ Strain, data = fibroblast_amplitude_resid))
```

The Box-Cox log-likelihood is highest at 0, which implies that a log-transform will be advantageous.

```{r}
boxcox(lm(log10(Amplitude_Mean) ~ Strain, data = fibroblast_amplitude_resid))
```

The Box-Cox is centered closer to 1. Plotting residuals.

```{r}
fibroblast_amplitude_resid$Amplitude_Resid_log10 = resid(lm(log10(Amplitude_Mean) ~ Strain, data = fibroblast_amplitude_resid))

ggplot(data = fibroblast_amplitude_resid, aes(x = Strain, y = Amplitude_Resid_log10, color = Strain, shape = Sex)) +
  geom_hline(yintercept = 0, col = "#999999") +
  geom_beeswarm(size = 2) + 
  scale_color_manual(values = founders_key$collaborative_cross_color_broman) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_shape_manual(values = c(16,23))
```

There still appears to be some heteroscedasticity, but to within the tolerance of a linear model. This looks good. Comparing AICs for different possible models.

```{r}
AICs = c()
AICs["AIC_inter"] = AIC(lm(log10(Amplitude_Mean) ~ Cohort + Strain + Sex + Strain:Sex, data = fibroblast_amplitude_mean))
AICs["AIC_nointer"] = AIC(lm(log10(Amplitude_Mean) ~ Cohort + Strain + Sex, data = fibroblast_amplitude_mean))
AICs["AIC_nosex"] = AIC(lm(log10(Amplitude_Mean) ~ Cohort + Strain, data = fibroblast_amplitude_mean))
AICs["AIC_nocohort"] = AIC(lm(log10(Amplitude_Mean) ~ Strain, data = fibroblast_amplitude_mean))
AICs["AIC_nostrain"] = AIC(lm(log10(Amplitude_Mean) ~ Cohort, data = fibroblast_amplitude_mean))

AICs

BICs = c()
BICs["BIC_inter"] = BIC(lm(log10(Amplitude_Mean) ~ Cohort + Strain + Sex + Strain:Sex, data = fibroblast_amplitude_mean))
BICs["BIC_nointer"] = BIC(lm(log10(Amplitude_Mean) ~ Cohort + Strain + Sex, data = fibroblast_amplitude_mean))
BICs["BIC_nosex"] = BIC(lm(log10(Amplitude_Mean) ~ Cohort + Strain, data = fibroblast_amplitude_mean))
BICs["BIC_nocohort"] = BIC(lm(log10(Amplitude_Mean) ~ Strain, data = fibroblast_amplitude_mean))
BICs["BIC_nostrain"] = BIC(lm(log10(Amplitude_Mean) ~ Cohort, data = fibroblast_amplitude_mean))

BICs
```

The best model for heritability according to both AIC and BIC contains additive covariates of cohort and strain, but not sex and strain:sex interaction factors. Plotting the residuals for this model.

```{r}
fibroblast_amplitude_resid$Amplitude_Resid_log10_CohortStrain = resid(lm(log10(Amplitude_Mean) ~ Cohort + Strain, data = fibroblast_amplitude_resid))

ggplot(data = fibroblast_amplitude_resid, aes(x = Strain, y = Amplitude_Resid_log10_CohortStrain, color = Strain, shape = Sex)) +
  geom_hline(yintercept = 0, col = "#999999") +
  geom_beeswarm(size = 2) + 
  scale_color_manual(values = founders_key$collaborative_cross_color_broman) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_shape_manual(values = c(16,23))
```

This model has a good fit. Showing ANOVA table for this model.

```{r}
anova(lm(log10(Amplitude_Mean) ~ Cohort + Strain, data = fibroblast_amplitude_resid))
```

This will be used for downstream heritability calculations. However, also fitting overall model with strain, sex, and their interactions:

```{r}
anova(lm(log10(Amplitude_Mean) ~ Cohort + Strain + Sex + Strain:Sex, data = fibroblast_amplitude_resid))
```

### Calculating heritability for fibroblast data

Calculating heritability using the ANOVA table from the best fitting model. Using the following equation:

$h^2=\displaystyle{{MS_{strain}}\over{MS_{strain} + (n_{mean} - 1) * MS_{resid}}}$

where $MS_{strain}$ is the mean square of the strain differences, $MS_{resid}$ is the mean square of the residuals, and $n_{mean}$ is the mean sample size of a strain.

```{r}
mean_n = nrow(fibroblast_amplitude_mean) / length(unique(fibroblast_amplitude_mean$Strain))
anova_amplitude = anova(lm(log10(Amplitude_Mean) ~ Cohort + Strain, data = fibroblast_amplitude_mean))
anova_amplitude_df = as.data.frame(anova_amplitude)
h2_amplitude = anova_amplitude_df["Strain","Mean Sq"] / (anova_amplitude_df["Strain","Mean Sq"] + ((mean_n - 1) * anova_amplitude_df["Residuals","Mean Sq"]))
h2_amplitude
```

### Plotting strain x sex means

Plotting the strain:sex information for these data.

```{r}

fibroblast_plotting_summary = fibroblast_amplitude_mean %>%
  group_by(Strain, Sex) %>%
  summarize(Strain_Sex_mean = mean(log10(Amplitude_Mean), na.rm = TRUE),
            Strain_Sex_sem = sd(log10(Amplitude_Mean), na.rm = TRUE) / length(which(!is.na(log10(Amplitude_Mean))))) %>%
  mutate(Strain_Sex = paste(Strain, gsub("^([MF]).*$", "\\1", Sex), sep = " "),
         Strain_Sex_backmean = 10^Strain_Sex_mean,
         Strain_Sex_ul = 10^(Strain_Sex_mean + Strain_Sex_sem), 
         Strain_Sex_ll = 10^(Strain_Sex_mean - Strain_Sex_sem))
fibroblast_plotting_summary = as.data.frame(fibroblast_plotting_summary)
fibroblast_plotting_summary$Strain_Sex = factor(fibroblast_plotting_summary$Strain_Sex,
                                                levels = paste(rep(founders_key$strain, each = 2), rep(c(" F", " M"), times = 8), sep = ""),
                                                ordered = TRUE)
fibroblast_plotting_summary$Strain = factor(fibroblast_plotting_summary$Strain,
                                            levels = founders_key$strain,
                                            ordered = TRUE)

founders_amplitude_strain_sex_plot = ggplot(data = fibroblast_plotting_summary, aes(x = Strain_Sex, y = Strain_Sex_backmean, color = Strain, group = Strain)) +
  geom_line(size = 1) +
  geom_point(size = 2) +
  geom_errorbar(aes(ymin = Strain_Sex_ll, ymax = Strain_Sex_ul), width = 0.5, size = 1) +
  scale_color_manual(values = founders_key$collaborative_cross_color_broman) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position="none") +
  xlab(NULL) +
  ylab("Amplitude (log-scale)") +
  scale_y_log10() +
  ggtitle("Fibroblast Rhythm: Amplitude")

founders_amplitude_strain_sex_plot
```

Saving plot

```{r}
saveRDS(founders_amplitude_strain_sex_plot, "./output/founders_amplitude_strain_sex_plot.RDS")
```

# Reproducibility Information

## Software Used

### R session and OS information

```{r sessionInfo}
circadian_prep_sessionInfo = sessionInfo()
circadian_prep_sessionInfo
saveRDS(circadian_prep_sessionInfo, paste(script_root_dir, "./output/circadian_prep_sessionInfo_",current_date,".RDS", sep = ""), compress="xz")
```

## Document Control

This document was prepared using RMarkdown in RStudio.
