---
title: "Mouse MGP VCF File: DO and CC Founders"
author: "Michael C. Saul (michael.saul [at] jax.org)"
date: "2018-10-25"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: false
      smooth_scroll: false
    number_sections: true
    fig_retina: true
    self_contained: true
    code_folding: show
    theme: flatly
    highlight: default
    df_print: paged
---

# Background

## Motivation

### Purpose of analysis

The purpose of this analysis is to reproducibly download the founder Sanger MGP SNP VCF files for release 1505 of the MGP in advance of compiling them into a single file for use in Diversity Outbred RNA-seq alignment.

# Analysis

## Downloads

### Founder Variants (SNPs)

Downloading and checking md5sums of the following SNP files:

File Name                                 | md5 Checksum
------------------------------------------|------------------------------------
129S1_SvImJ.mgp.v5.snps.dbSNP142.vcf.gz   | `935ff15ff2ab878d961121c303ced19d` 
A_J.mgp.v5.snps.dbSNP142.vcf.gz           | `fb53f6a71c607c59c61bd2fa1f0efa49` 
CAST_EiJ.mgp.v5.snps.dbSNP142.vcf.gz      | `7180ebf308a5ffb44e733fcb6b85a7d0` 
NOD_ShiLtJ.mgp.v5.snps.dbSNP142.vcf.gz    | `fcf8319a19e7d3720ab64bec705671c4`
NZO_HlLtJ.mgp.v5.snps.dbSNP142.vcf.gz     | `0f9ac892d52e2cf1aef09b4060c0411e`
PWK_PhJ.mgp.v5.snps.dbSNP142.vcf.gz       | `62da476180985068c440b13591ead0a6`
WSB_EiJ.mgp.v5.snps.dbSNP142.vcf.gz       | `ea93453a6c94ad781ce8c155c4538966`

```{sh}
# Shell script
echo "Beginning release 1505 founder SNP download"
curl -s -O ftp://ftp-mouse.sanger.ac.uk/REL-1505-SNPs_Indels/strain_specific_vcfs/129S1_SvImJ.mgp.v5.snps.dbSNP142.vcf.gz 
curl -s -O ftp://ftp-mouse.sanger.ac.uk/REL-1505-SNPs_Indels/strain_specific_vcfs/A_J.mgp.v5.snps.dbSNP142.vcf.gz
curl -s -O ftp://ftp-mouse.sanger.ac.uk/REL-1505-SNPs_Indels/strain_specific_vcfs/CAST_EiJ.mgp.v5.snps.dbSNP142.vcf.gz
curl -s -O ftp://ftp-mouse.sanger.ac.uk/REL-1505-SNPs_Indels/strain_specific_vcfs/NOD_ShiLtJ.mgp.v5.snps.dbSNP142.vcf.gz
curl -s -O ftp://ftp-mouse.sanger.ac.uk/REL-1505-SNPs_Indels/strain_specific_vcfs/NZO_HlLtJ.mgp.v5.snps.dbSNP142.vcf.gz 
curl -s -O ftp://ftp-mouse.sanger.ac.uk/REL-1505-SNPs_Indels/strain_specific_vcfs/PWK_PhJ.mgp.v5.snps.dbSNP142.vcf.gz
curl -s -O ftp://ftp-mouse.sanger.ac.uk/REL-1505-SNPs_Indels/strain_specific_vcfs/WSB_EiJ.mgp.v5.snps.dbSNP142.vcf.gz 
echo "Finished release 1505 founder SNP download"
```

Checking md5 checksums of these files

```{r}
# R code
library("tools")
paste("The md5 checksum for the downloaded 129S1_SvImJ.mgp.v5.snps.dbSNP142.vcf.gz",ifelse(md5sum("129S1_SvImJ.mgp.v5.snps.dbSNP142.vcf.gz") == "935ff15ff2ab878d961121c303ced19d", " does ", " does not "), "agree with the reference: 935ff15ff2ab878d961121c303ced19d", sep = "")
paste("The md5 checksum for the downloaded A_J.mgp.v5.snps.dbSNP142.vcf.gz",ifelse(md5sum("A_J.mgp.v5.snps.dbSNP142.vcf.gz") == "fb53f6a71c607c59c61bd2fa1f0efa49", " does ", " does not "), "agree with the reference: fb53f6a71c607c59c61bd2fa1f0efa49", sep = "")
paste("The md5 checksum for the downloaded CAST_EiJ.mgp.v5.snps.dbSNP142.vcf.gz",ifelse(md5sum("CAST_EiJ.mgp.v5.snps.dbSNP142.vcf.gz") == "7180ebf308a5ffb44e733fcb6b85a7d0", " does ", " does not "), "agree with the reference: 7180ebf308a5ffb44e733fcb6b85a7d0", sep = "")
paste("The md5 checksum for the downloaded NOD_ShiLtJ.mgp.v5.snps.dbSNP142.vcf.gz",ifelse(md5sum("NOD_ShiLtJ.mgp.v5.snps.dbSNP142.vcf.gz") == "fcf8319a19e7d3720ab64bec705671c4", " does ", " does not "), "agree with the reference: fcf8319a19e7d3720ab64bec705671c4", sep = "")
paste("The md5 checksum for the downloaded NZO_HlLtJ.mgp.v5.snps.dbSNP142.vcf.gz",ifelse(md5sum("NZO_HlLtJ.mgp.v5.snps.dbSNP142.vcf.gz") == "0f9ac892d52e2cf1aef09b4060c0411e", " does ", " does not "), "agree with the reference: 0f9ac892d52e2cf1aef09b4060c0411e", sep = "")
paste("The md5 checksum for the downloaded PWK_PhJ.mgp.v5.snps.dbSNP142.vcf.gz",ifelse(md5sum("PWK_PhJ.mgp.v5.snps.dbSNP142.vcf.gz") == "62da476180985068c440b13591ead0a6", " does ", " does not "), "agree with the reference: 62da476180985068c440b13591ead0a6", sep = "")
paste("The md5 checksum for the downloaded WSB_EiJ.mgp.v5.snps.dbSNP142.vcf.gz",ifelse(md5sum("WSB_EiJ.mgp.v5.snps.dbSNP142.vcf.gz") == "ea93453a6c94ad781ce8c155c4538966", " does ", " does not "), "agree with the reference: ea93453a6c94ad781ce8c155c4538966", sep = "")
```

### Founder Variants (INDELs)

Downloading and checking md5sums of the following INDEL files:

File Name                                         | md5 Checksum
--------------------------------------------------|------------------------------------
129S1_SvImJ.mgp.v5.indels.dbSNP142.normed.vcf.gz  | `faa52363b8c6b25016ad87a1b8b3ae06` 
A_J.mgp.v5.indels.dbSNP142.normed.vcf.gz          | `14c137fef26739fd914fae181491812c` 
CAST_EiJ.mgp.v5.indels.dbSNP142.normed.vcf.gz     | `a92681cf1dd5108ac36e17b5e9313a3f` 
NOD_ShiLtJ.mgp.v5.indels.dbSNP142.normed.vcf.gz   | `55b4d07c7b04af2d92ce59b056d9f447`
NZO_HlLtJ.mgp.v5.indels.dbSNP142.normed.vcf.gz    | `07f4399639fd79170cf95d0a0f776aad`
PWK_PhJ.mgp.v5.indels.dbSNP142.normed.vcf.gz      | `ac9c0aec191456aa8e75f670c1f82d0c`
WSB_EiJ.mgp.v5.indels.dbSNP142.normed.vcf.gz      | `dd95f46dcbdf0bc4a7d8f65c47088c19`

```{sh}
# Shell script
echo "Beginning release 1505 founder INDEL download"
curl -s -O ftp://ftp-mouse.sanger.ac.uk/REL-1505-SNPs_Indels/strain_specific_vcfs/129S1_SvImJ.mgp.v5.indels.dbSNP142.normed.vcf.gz
curl -s -O ftp://ftp-mouse.sanger.ac.uk/REL-1505-SNPs_Indels/strain_specific_vcfs/A_J.mgp.v5.indels.dbSNP142.normed.vcf.gz
curl -s -O ftp://ftp-mouse.sanger.ac.uk/REL-1505-SNPs_Indels/strain_specific_vcfs/CAST_EiJ.mgp.v5.indels.dbSNP142.normed.vcf.gz
curl -s -O ftp://ftp-mouse.sanger.ac.uk/REL-1505-SNPs_Indels/strain_specific_vcfs/NOD_ShiLtJ.mgp.v5.indels.dbSNP142.normed.vcf.gz
curl -s -O ftp://ftp-mouse.sanger.ac.uk/REL-1505-SNPs_Indels/strain_specific_vcfs/NZO_HlLtJ.mgp.v5.indels.dbSNP142.normed.vcf.gz
curl -s -O ftp://ftp-mouse.sanger.ac.uk/REL-1505-SNPs_Indels/strain_specific_vcfs/PWK_PhJ.mgp.v5.indels.dbSNP142.normed.vcf.gz
curl -s -O ftp://ftp-mouse.sanger.ac.uk/REL-1505-SNPs_Indels/strain_specific_vcfs/WSB_EiJ.mgp.v5.indels.dbSNP142.normed.vcf.gz 
echo "Finished release 1505 founder INDEL download"
```

Checking md5 checksums of these files

```{r}
# R code
paste("The md5 checksum for the downloaded 129S1_SvImJ.mgp.v5.indels.dbSNP142.normed.vcf.gz",ifelse(md5sum("129S1_SvImJ.mgp.v5.indels.dbSNP142.normed.vcf.gz") == "faa52363b8c6b25016ad87a1b8b3ae06", " does ", " does not "), "agree with the reference: faa52363b8c6b25016ad87a1b8b3ae06", sep = "")
paste("The md5 checksum for the downloaded A_J.mgp.v5.indels.dbSNP142.normed.vcf.gz",ifelse(md5sum("A_J.mgp.v5.indels.dbSNP142.normed.vcf.gz") == "14c137fef26739fd914fae181491812c", " does ", " does not "), "agree with the reference: 14c137fef26739fd914fae181491812c", sep = "")
paste("The md5 checksum for the downloaded CAST_EiJ.mgp.v5.indels.dbSNP142.normed.vcf.gz",ifelse(md5sum("CAST_EiJ.mgp.v5.indels.dbSNP142.normed.vcf.gz") == "a92681cf1dd5108ac36e17b5e9313a3f", " does ", " does not "), "agree with the reference: a92681cf1dd5108ac36e17b5e9313a3f", sep = "")
paste("The md5 checksum for the downloaded NOD_ShiLtJ.mgp.v5.indels.dbSNP142.normed.vcf.gz",ifelse(md5sum("NOD_ShiLtJ.mgp.v5.indels.dbSNP142.normed.vcf.gz") == "55b4d07c7b04af2d92ce59b056d9f447", " does ", " does not "), "agree with the reference: 55b4d07c7b04af2d92ce59b056d9f447", sep = "")
paste("The md5 checksum for the downloaded NZO_HlLtJ.mgp.v5.indels.dbSNP142.normed.vcf.gz",ifelse(md5sum("NZO_HlLtJ.mgp.v5.indels.dbSNP142.normed.vcf.gz") == "07f4399639fd79170cf95d0a0f776aad", " does ", " does not "), "agree with the reference: 07f4399639fd79170cf95d0a0f776aad", sep = "")
paste("The md5 checksum for the downloaded PWK_PhJ.mgp.v5.indels.dbSNP142.normed.vcf.gz",ifelse(md5sum("PWK_PhJ.mgp.v5.indels.dbSNP142.normed.vcf.gz") == "ac9c0aec191456aa8e75f670c1f82d0c", " does ", " does not "), "agree with the reference: ac9c0aec191456aa8e75f670c1f82d0c", sep = "")
paste("The md5 checksum for the downloaded WSB_EiJ.mgp.v5.indels.dbSNP142.normed.vcf.gz",ifelse(md5sum("WSB_EiJ.mgp.v5.indels.dbSNP142.normed.vcf.gz") == "dd95f46dcbdf0bc4a7d8f65c47088c19", " does ", " does not "), "agree with the reference: dd95f46dcbdf0bc4a7d8f65c47088c19", sep = "")
```

### Genome Files

Downloading the soft masked primary assembly of the GRCm38 genome and the Ensembl v94 transcriptome annotation. Ensembl does not list md5 checksums, but they do list conventional checksums.

File Name                                           | Checksum
----------------------------------------------------|---------------------
Mus_musculus.GRCm38.dna_sm.primary_assembly.fa.gz   | `04137 841792` 
Mus_musculus.GRCm38.94.gtf.gz                       | `15062 28709`

```{sh}
# Shell script
echo "Beginning GRCm38 genome and Ensembl v94 annotation download"
curl -s -O ftp://ftp.ensembl.org/pub/release-94/fasta/mus_musculus/dna/Mus_musculus.GRCm38.dna_sm.primary_assembly.fa.gz
curl -s -O ftp://ftp.ensembl.org/pub/release-94/gtf/mus_musculus/Mus_musculus.GRCm38.94.gtf.gz
echo "Finished GRCm38 genome and Ensembl v94 annotation download"
```

Checking the checksums of these files.

```{r}
# R code
paste("The checksum for the downloaded Mus_musculus.GRCm38.dna_sm.primary_assembly.fa.gz",
      ifelse(system("cksum -o 1 Mus_musculus.GRCm38.dna_sm.primary_assembly.fa.gz", intern = TRUE) == "4137 841792 Mus_musculus.GRCm38.dna_sm.primary_assembly.fa.gz", 
             " does ", " does not "), "agree with the reference: 04137 841792", sep = "")
paste("The checksum for the downloaded Mus_musculus.GRCm38.94.gtf.gz",
      ifelse(system("cksum -o 1 Mus_musculus.GRCm38.94.gtf.gz", intern = TRUE) == "15062 28709 Mus_musculus.GRCm38.94.gtf.gz", 
             " does ", " does not "), "agree with the reference: 15062 28709", sep = "")
```

Getting the md5 checksums for these files.

```{r}
# R code
paste("The md5 checksum for the downloaded Mus_musculus.GRCm38.dna_sm.primary_assembly.fa.gz is: ",
      md5sum("Mus_musculus.GRCm38.dna_sm.primary_assembly.fa.gz"), sep = "")
paste("The checksum for the downloaded Mus_musculus.GRCm38.94.gtf.gz is: ",
      md5sum("Mus_musculus.GRCm38.94.gtf.gz"), sep = "")
```

## Reproducibility Information

### R session and OS information

```{r}
# R code
founders_mgp_prep_sessionInfo = sessionInfo()
founders_mgp_prep_sessionInfo
```

### Document Control

This document was prepared using RMarkdown in RStudio.

